---
title: "Pelvic mesh recall signal"
author: "Curtis Murray"
mail: "curtis.murray@adelaide.edu.au"
linkedin: "curtiswmurray"
twitter: "curtis_w_murray"
github: "curtis-murray"
home: "curtismurray.tech"
#logo: "logo_gallery.png"
output:
  bookdown::html_document2:
    keep_tex: true
    theme: "lumen"
#    template: resources/template.html
    css: resources/style.css
    toc: TRUE
    code_folding: "show"
    number_sections: FALSE
    toc_depth: 2
    toc_float: TRUE
    self_contained: FALSE
#    pandoc_args: --mathjax=https://cdnjs.cloudflare.com/ajax/libs/mathjax/2.7.5/latest.js?config=TeX-MML-AM_HTMLorMML.js
#    self_contained: FALSE
---

```{css zoom-lib-src, echo = FALSE}
script src = "https://ajax.googleapis.com/ajax/libs/jquery/3.4.1/jquery.min.js"
```

```{js zoom-jquery, echo = FALSE}
$(document).ready(function() {
$('body').prepend('<div class=\"zoomDiv\"><img src=\"\" class=\"zoomImg\"></div>');
// onClick function for all plots (img's)
$('img:not(.zoomImg)').click(function() {
$('.zoomImg').attr('src', $(this).attr('src')).css({width: '100%'});
$('.zoomDiv').css({opacity: '1', width: 'auto', border: '1px solid white', borderRadius: '5px', position: 'fixed', top: '50%', left: '50%', marginRight: '-50%', transform: 'translate(-50%, -50%)', boxShadow: '0px 0px 50px #888888', zIndex: '50', overflow: 'auto', maxHeight: '100%'});
});
// onClick function for zoomImg
$('img.zoomImg').click(function() {
$('.zoomDiv').css({opacity: '0', width: '0%'}); 

});

$('.my_card').on('click',function(e) {
$(this).addClass('active');
e.stopPropagation()
});
$(document).on("click", function(e) {
if ($(e.target).is(".my_card") === false) {
$(".my_card").removeClass("active");
}
});
});
```


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
library(epuRate)
library(rmarkdown)
library(tidyverse)
library(lubridate)
library(mds)
library(mdsstat)
```

```{r}
df <- read_csv("data/all_reports/all_reports_mached_manufacturer_df.csv")
```

Testing for statistical significance between pelvic mesh reports and other reports to identify a potential signal that could have been used to recall pelvic mesh devices prior to the 2018 recall.

## Fisher test of pain

### Pain in pelvic mesh devices and hernia mesh devices

First we look at mentions of pain in pelvic mesh devices and hernia mesh devices. We construct a contingency table counting the number of reports that have mentions of the word `pain' for both devices.


```{r}
# Fisher test that the odds ratio between pain levels in pelvic reports and other reports is equal to 1.

# -------
# Pelvic vs hernia
mesh_filtered <- df %>% 
  mutate(`GMDN term` = str_to_lower(`GMDN term`)) %>%
  filter(str_detect(`GMDN term`, "mesh")) %>% 
  group_by(`GMDN term`) %>% 
  mutate(hernia = str_detect(`Event description`, "(?i)hernia"),
         pelvic = str_detect(`Event description`, "(?i)((pelvi)|(urina)|(incont)|(vagina)|(uterus)|(anterior)|(posterior))")) %>% 
  filter(hernia | pelvic) %>% 
  filter(!str_detect(`Event description`,"Journal Article")) %>% 
  filter(`Report date` < "2018-01-01")

mesh_pain <- mesh_filtered %>% 
  mutate(pain = str_detect(`Event description`, "(?i)pain")) %>% 
  group_by(hernia, pelvic, pain) %>% 
  summarise(count = n())


pain_table <- mesh_pain %>% 
  ungroup() %>% 
  mutate(type = ifelse(pelvic, "pelvic","hernia"),
         pain = ifelse(pain, "pain","no pain")) %>% 
  select(type,pain,count) %>% 
  pivot_wider(values_from = count, names_from = pain) %>% 
  select(type, pain, `no pain`) %>% 
  as.data.frame()

rownames(pain_table) <- pain_table$type
pain_table <- pain_table[,2:3]
pain_table
```

We use Fisher's exact test to test the null hypothesis that the odds ratio is equal to one for

$$ Odds Ratio = \frac{Pain_P / NoPain_P}{Pain_H / NoPain_H}$$

```{r}
fisher.test(pain_table)
```

We reject the null hypothesis that the odds ratio is equal to one $\left(p = 3e-12\right)$, with a $95\%$ confidence interval of $(7.22, 91.7)$. The odds ratio can be interpreted as how many times greater the ratio of pain to no pain documents is between pelvic mesh and hernia mesh devices is.

### Pain in pelvic mesh devices and all other devices

```{r}
# --------- 
# Pelvic against all other devices

mesh_filtered <- df %>% 
  mutate(`GMDN term` = str_to_lower(`GMDN term`)) %>%
  mutate(
    pelvic = (str_detect(`Event description`, "(?i)((pelvi)|(urina)|(incont)|(vagina)|(uterus)|(anterior)|(posterior))") & 
                str_detect(`GMDN term`, "mesh"))
  ) %>% 
  mutate(other = !pelvic) %>% 
  filter(`Report date` < "2018-01-01")

mesh_pain <- mesh_filtered %>% 
  mutate(pain = str_detect(`Event description`, "(?i)pain")) %>% 
  group_by(other, pelvic, pain) %>% 
  summarise(count = n())

pain_table <- mesh_pain %>% 
  ungroup() %>% 
  mutate(type = ifelse(pelvic, "pelvic","other"),
         pain = ifelse(pain, "pain","no pain")) %>% 
  select(type,pain,count) %>% 
  pivot_wider(values_from = count, names_from = pain) %>% 
  select(type, pain, `no pain`) %>% 
  as.data.frame()

rownames(pain_table) <- pain_table$type
pain_table <- pain_table[,2:3]

pain_table
```

```{r}
fisher.test(pain_table)
```

## Fisher test of pain topic
### Pain topic in pelvic mesh devices and hernia mesh devices

We can do the same thing but looking at a `pain' topic.

As a start, I've picked the below topic, and set a threshold of $0.05$ to indicate its presence in a document.

```{r}
library(ggwordcloud)

tidy_topics <- read_csv("data/hSBM/Clean/tidy_topics.csv")
tidy_topics_docs <- read_csv("data/hSBM/Clean/tidy_topics_docs.csv")

pain_topic_1 <- tidy_topics %>% filter(word == "pain", Level == 1) %>% 
  pull(topic)

wc_data <- tidy_topics %>% 
  filter(Level == 1, topic == pain_topic_1)

ggwordcloud(words = wc_data$word, freq = wc_data$p)
```

```{r}
mesh_filtered <- df %>% 
  mutate(`GMDN term` = str_to_lower(`GMDN term`)) %>%
  filter(str_detect(`GMDN term`, "mesh")) %>% 
  group_by(`GMDN term`) %>% 
  mutate(hernia = str_detect(`Event description`, "(?i)hernia"),
         pelvic = str_detect(`Event description`, "(?i)((pelvi)|(urina)|(incont)|(vagina)|(uterus)|(anterior)|(posterior))")) %>% 
  filter(hernia | pelvic) %>% 
  filter(!str_detect(`Event description`,"Journal Article"))

mesh_pain <- mesh_filtered %>% 
  ungroup() %>% 
  left_join(tidy_topics_docs %>% 
              filter(Level == 1, topic == pain_topic_1),
            by = c("Report number" = "document")) %>% 
  select(hernia, pelvic, p, `Event description`) %>% 
  mutate(pain = ifelse(p >= 0.3, "pain threshold", "no pain threshold")) %>% 
  mutate(pain = ifelse(is.na(p), "no pain threshold",pain)) %>% 
  group_by(hernia, pelvic, pain) %>% 
  summarise(count = n())

pain_table <- mesh_pain %>% 
  ungroup() %>% 
  mutate(type = ifelse(pelvic, "pelvic","hernia")) %>% 
  select(type,pain,count) %>% 
  pivot_wider(values_from = count, names_from = pain) %>% 
  select(type, `pain threshold`, `no pain threshold`) %>% 
  as.data.frame()


rownames(pain_table) <- pain_table$type
pain_table <- pain_table[,2:3]
pain_table

```

```{r}
fisher.test(pain_table)
```

### Pain topic in pelvic mesh devices and all other devices

```{r}
mesh_filtered <- df %>% 
  mutate(`GMDN term` = str_to_lower(`GMDN term`)) %>%
  mutate(
    pelvic = (str_detect(`Event description`, "(?i)((pelvi)|(urina)|(incont)|(vagina)|(uterus)|(anterior)|(posterior))") & 
                str_detect(`GMDN term`, "mesh"))
  ) %>% 
  mutate(other = !pelvic) %>% 
  filter(`Report date` < "2018-01-01")

mesh_pain <- mesh_filtered %>% 
  ungroup() %>% 
  left_join(tidy_topics_docs %>% 
              filter(Level == 1, topic == pain_topic_1),
            by = c("Report number" = "document")) %>% 
  select(other, pelvic, p, `Event description`) %>% 
  mutate(pain = ifelse(p >= 0.3, "pain threshold", "no pain threshold")) %>% 
  mutate(pain = ifelse(is.na(p), "no pain threshold",pain)) %>% 
  group_by(other, pelvic, pain) %>% 
  summarise(count = n())

pain_table <- mesh_pain %>% 
  ungroup() %>% 
  mutate(type = ifelse(pelvic, "pelvic","other")) %>% 
  select(type,pain,count) %>% 
  pivot_wider(values_from = count, names_from = pain) %>% 
  select(type, `pain threshold`, `no pain threshold`) %>% 
  as.data.frame()


rownames(pain_table) <- pain_table$type
pain_table <- pain_table[,2:3]
pain_table

```

```{r}
fisher.test(pain_table)
```

## Disproportionality analysis

```{r}
# Save a list of algorithms to run
x <- list(prr=list(),
            xbar=list(),
            xbar=list(ts_event=c(Rate="rate"), we_rule=2),
            poisson_rare=list(p_rate=0.3),
          bcpnn = list())
algos <- define_algos(x)
```

### Pain in pelvic mesh devices and hernia mesh devices

```{r}

unit = "years"

mesh_filtered <- df %>% 
  mutate(`GMDN term` = str_to_lower(`GMDN term`)) %>%
  filter(str_detect(`GMDN term`, "mesh")) %>% 
  group_by(`GMDN term`) %>% 
  mutate(hernia = str_detect(`Event description`, "(?i)hernia"),
         pelvic = str_detect(`Event description`, "(?i)((pelvi)|(urina)|(incont)|(vagina)|(uterus)|(anterior)|(posterior))")) %>% 
  filter(hernia | pelvic) %>% 
  filter(!str_detect(`Event description`,"Journal Article")) %>% 
  filter(`Report date` < "2018-01-01") %>% 
  ungroup() %>% 
  mutate(Date = floor_date(`Report date`, unit = unit))

mesh_pain <- mesh_filtered %>% 
  mutate(pain = str_detect(`Event description`, "(?i)pain")) %>% 
  group_by(hernia, pelvic, pain, Date) %>% 
  summarise(count = n())


pain_table <- mesh_pain %>% 
  ungroup() %>% 
  mutate(type = ifelse(pelvic, "pelvic","hernia"),
         pain = ifelse(pain, "pain","no pain")) %>% 
  select(Date,type,pain,count) %>% 
  pivot_wider(values_from = count, names_from = pain) %>% 
  as.data.frame() %>% 
  arrange(Date) %>% 
  bind_rows(
    tibble(Date = seq.Date(from = min(mesh_pain$Date), to = max(mesh_pain$Date), by = unit)) %>% 
      mutate(type = "pelvic", `no pain` = 0, `pain` = 0)
  ) %>% 
  bind_rows(
    tibble(Date = seq.Date(from = min(mesh_pain$Date), to = max(mesh_pain$Date), by = unit)) %>% 
      mutate(type = "hernia", `no pain` = 0, `pain` = 0)
  ) %>% 
  group_by(Date, type) %>% 
  summarise(pain = sum(pain, na.rm = T),
            `no pain` = sum(`no pain`, na.rm = T)) %>% 
  pivot_wider(names_from = type, values_from = c("pain", "no pain")) %>% 
  ungroup() %>% 
  transmute(time = Date, nA = pain_pelvic, nB = `no pain_pelvic`, nC = pain_hernia, nD = `no pain_hernia`,event = nA) %>% 
  as.data.frame() 
pain_table
run_algos(pain_table, algos)

```

### Pain topic in pelvic mesh devices and hernial mesh devices

```{r}
#----- Pain topic hernia

mesh_filtered <- df %>% 
  mutate(`GMDN term` = str_to_lower(`GMDN term`)) %>%
  filter(str_detect(`GMDN term`, "mesh")) %>% 
  group_by(`GMDN term`) %>% 
  mutate(hernia = str_detect(`Event description`, "(?i)hernia"),
         pelvic = str_detect(`Event description`, "(?i)((pelvi)|(urina)|(incont)|(vagina)|(uterus)|(anterior)|(posterior))")) %>% 
  filter(hernia | pelvic) %>% 
  filter(!str_detect(`Event description`,"Journal Article")) %>% 
  filter(`Report date` < "2018-01-01") %>% 
  mutate(Date = floor_date(`Report date`, unit = unit))

mesh_pain <- mesh_filtered %>% 
  ungroup() %>% 
  left_join(tidy_topics_docs %>% 
              filter(Level == 1, topic == pain_topic_1),
            by = c("Report number" = "document")) %>% 
  select(hernia, pelvic, p, `Event description`, Date) %>% 
  mutate(pain = ifelse(p >= 0.05, "pain threshold", "no pain threshold")) %>% 
  mutate(pain = ifelse(is.na(p), "no pain threshold",pain)) %>% 
  group_by(hernia, pelvic, pain, Date) %>% 
  summarise(count = n()) 

pain_table <- mesh_pain %>% 
  ungroup() %>% 
  mutate(type = ifelse(pelvic, "pelvic","hernia")) %>% 
  select(Date,type,pain,count) %>% 
  pivot_wider(values_from = count, names_from = pain) %>% 
  as.data.frame() %>% 
  arrange(Date) %>% 
  bind_rows(
    tibble(Date = seq.Date(from = min(mesh_pain$Date), to = max(mesh_pain$Date), by = unit)) %>% 
      mutate(type = "pelvic", `no pain` = 0, `pain` = 0)
  ) %>% 
  bind_rows(
    tibble(Date = seq.Date(from = min(mesh_pain$Date), to = max(mesh_pain$Date), by = unit)) %>% 
      mutate(type = "hernia", `no pain` = 0, `pain` = 0)
  ) %>% 
  group_by(Date, type) %>% 
  summarise(`pain threshold` = sum(`pain threshold`, na.rm = T),
            `no pain threshold` = sum(`no pain threshold`, na.rm = T)) %>% 
  pivot_wider(names_from = type, values_from = c("pain threshold", "no pain threshold")) %>% 
  ungroup() %>% 
  transmute(time = Date, nA = `pain threshold_pelvic`, nB = `no pain threshold_pelvic`, nC = `pain threshold_hernia`, nD = `no pain threshold_hernia`,event = nA) %>% 
  as.data.frame()

pain_table
run_algos(pain_table, algos)
```

### Pain topic in pelvic mesh devices and all other devices

```{r}
#----- Pain topic other

mesh_filtered <- df %>% 
  mutate(`GMDN term` = str_to_lower(`GMDN term`)) %>%
  mutate(
    pelvic = (str_detect(`Event description`, "(?i)((pelvi)|(urina)|(incont)|(vagina)|(uterus)|(anterior)|(posterior))") & 
                str_detect(`GMDN term`, "mesh"))
  ) %>% 
  mutate(other = !pelvic) %>% 
  ungroup() %>% 
  mutate(Date = floor_date(`Report date`, unit = unit)) %>% 
  filter(Date < "2018-01-01")

mesh_pain <- mesh_filtered %>% 
  ungroup() %>% 
  left_join(tidy_topics_docs %>% 
              filter(Level == 1, topic == pain_topic_1),
            by = c("Report number" = "document")) %>% 
  select(other, pelvic, p, `Event description`, Date) %>% 
  mutate(pain = ifelse(p >= 0.05, "pain threshold", "no pain threshold")) %>% 
  mutate(pain = ifelse(is.na(p), "no pain threshold",pain)) %>% 
  group_by(other, pelvic, pain, Date) %>% 
  summarise(count = n()) 

pain_table <- mesh_pain %>% 
  ungroup() %>% 
  mutate(type = ifelse(pelvic, "pelvic","other")) %>% 
  select(Date,type,pain,count) %>% 
  pivot_wider(values_from = count, names_from = pain) %>% 
  arrange(Date) %>% 
  bind_rows(
    tibble(Date = seq.Date(from = min(mesh_pain$Date), to = max(mesh_pain$Date), by = unit)) %>% 
      mutate(type = "pelvic", `no pain threshold` = 0, `pain threshold` = 0)
  ) %>% 
  bind_rows(
    tibble(Date = seq.Date(from = min(mesh_pain$Date), to = max(mesh_pain$Date), by = unit)) %>% 
      mutate(type = "other", `no pain threshold` = 0, `pain threshold` = 0)
  ) %>% 
  group_by(Date, type) %>% 
  summarise(`pain threshold`= sum(`pain threshold`, na.rm = T),
            `no pain threshold` = sum(`no pain threshold`, na.rm = T)) %>% 
  pivot_wider(names_from = type, values_from = c("pain threshold", "no pain threshold")) %>% 
  ungroup() %>% 
  transmute(time = Date, nA = `pain threshold_pelvic`, nB = `no pain threshold_pelvic`, nC = `pain threshold_other`, nD = `no pain threshold_other`,event = nA) %>% 
  as.data.frame()

pain_table
run_algos(pain_table, algos)
```


### Pain topic in hernia mesh devices and all other devices

```{r}
#----- Pain topic other

mesh_filtered <- df %>% 
  mutate(`GMDN term` = str_to_lower(`GMDN term`)) %>%
  mutate(
    hernia = (str_detect(`Event description`, "(?i)hernia") & 
                str_detect(`GMDN term`, "mesh"))
  ) %>% 
  mutate(other = !hernia) %>% 
  ungroup() %>% 
  mutate(Date = floor_date(`Report date`, unit = unit)) %>% 
  filter(Date < "2018-01-01")

mesh_pain <- mesh_filtered %>% 
  ungroup() %>% 
  left_join(tidy_topics_docs %>% 
              filter(Level == 1, topic == pain_topic_1),
            by = c("Report number" = "document")) %>% 
  select(other, hernia, p, `Event description`, Date) %>% 
  mutate(pain = ifelse(p >= 0.05, "pain threshold", "no pain threshold")) %>% 
  mutate(pain = ifelse(is.na(p), "no pain threshold",pain)) %>% 
  group_by(other, hernia, pain, Date) %>% 
  summarise(count = n()) 

pain_table <- mesh_pain %>% 
  ungroup() %>% 
  mutate(type = ifelse(hernia, "hernia","other")) %>% 
  select(Date,type,pain,count) %>% 
  pivot_wider(values_from = count, names_from = pain) %>% 
  arrange(Date) %>% 
  bind_rows(
    tibble(Date = seq.Date(from = min(mesh_pain$Date), to = max(mesh_pain$Date), by = unit)) %>% 
      mutate(type = "hernia", `no pain threshold` = 0, `pain threshold` = 0)
  ) %>% 
  bind_rows(
    tibble(Date = seq.Date(from = min(mesh_pain$Date), to = max(mesh_pain$Date), by = unit)) %>% 
      mutate(type = "other", `no pain threshold` = 0, `pain threshold` = 0)
  ) %>% 
  group_by(Date, type) %>% 
  summarise(`pain threshold`= sum(`pain threshold`, na.rm = T),
            `no pain threshold` = sum(`no pain threshold`, na.rm = T)) %>% 
  pivot_wider(names_from = type, values_from = c("pain threshold", "no pain threshold")) %>% 
  ungroup() %>% 
  transmute(time = Date, nA = `pain threshold_hernia`, nB = `no pain threshold_hernia`, nC = `pain threshold_other`, nD = `no pain threshold_other`,event = nA) %>% 
  as.data.frame()

pain_table
run_algos(pain_table, algos)
```